---
title: "Studio 11"
author: "Mitchell Howard"
date: "12/2/2021"
output:
  word_document: default
  html_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(error = TRUE)

```

```{r}
library(dplyr)
library(tidyr)
library(stringr)
library(tidytext)

install.packages("dplyr")
install.packages("tidyr")
install.packages("stringr")
install.packages("tidytext")
install.packages("wordcloud")
```

```{r}

debatesdata <- read.csv("~/Downloads/primary_debates.csv", stringsAsFactors = FALSE)

candidates <- c('Bush', 'Carson', 'Chafee', 'Christie', 'Clinton', 'Cruz', 'Fiorina', 'Gilmore', 'Graham', 'Huckabee', 'Jindal', 'Kasich', 'O\'Malley', 'Pataki', 'Paul', 'Perry', 'Rubio', 'Sanders', 'Santorum', 'Trump', 'Walker', 'Webb')

debates <- debatesdata %>% filter(Speaker %in% candidates)

debates %>% count(debates$Speaker)


```

How many speakers are there? 

Asnwer: There are 22 speakers 

How many parties are there?

Answer: 3 parties

```{r}
debate_text <- debates %>% select(Party, Text)

debate_words <- debate_text %>% unnest_tokens(word, Text)


```


```{r}
debate_words %>% count(word, sort = TRUE) 

word_count <- debate_words %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE) 

```


Question #1: What are the five most common (non-stop) words said by the candidates?

Answer: People, Country, America, President, American


```{r}
library(ggplot2)
install.packages("ggplot2")
head(word_count, 20) %>% ggplot(aes(x = reorder(word, n), y = n)) + 
  geom_bar(stat = "identity") + coord_flip()



```

```{r wordcloud}

library(wordcloud)
install.packages("wordcloud")


word_count %>% with(wordcloud(word, n, max.words = 100))
word_count %>% with(wordcloud(word, n, max.words = 100, random.order = FALSE))
word_count %>% with(wordcloud(word, n, max.words = 100, random.order = FALSE, colors = 'darkcyan'))
word_count %>% with(wordcloud(word, n, max.words = 100, random.order = FALSE, colors = brewer.pal(8, "Dark2")))

```


```{r Democrats vs Republicans}

library(gridExtra)
install.packages("gridExtra")
word_count_by_party <- debate_words %>% anti_join(stop_words) %>% count(Party, word, sort = TRUE)
demtop15 <- word_count_by_party %>% group_by(Party) %>% filter(Party ==  "Democratic")
head(demtop15, 15) %>% ggplot(aes(x = reorder(word, n), y = n)) + 
  geom_bar(stat = "identity") + coord_flip()

reptop15 <-  word_count_by_party %>% group_by(Party) %>% filter(Party ==  "Republican")
head(reptop15, 15) %>% ggplot(aes(x = reorder(word, n), y = n)) + 
  geom_bar(stat = "identity") + coord_flip()


```

What words are in the top 15 for Democrats but not for Republicans? How about vice-versa?

Answer: Democrats said "health" "care" "wall" "street" "secretary" "world" "senator" and more while republicans talked more about "isis" "tax" "government" "clinton "hillary" "deal" "money" to name a few


```{r tf idf}
debates_tf_idf_top <- word_count_by_party %>%
  bind_tf_idf(word, Party, n) %>% 
  arrange(desc(tf_idf)) %>% 
  group_by(Party) %>%
  top_n(15)


ggplot(debates_tf_idf_top, aes(x = reorder(word, tf_idf), y = tf_idf, fill = Party)) +
  geom_bar(alpha = 0.8, stat = "identity", show.legend = FALSE) +
  facet_wrap(~Party, ncol = 2, scales = "free") + coord_flip() + 
  scale_fill_manual(values = c("Democratic" = "blue", "Republican" = "red"))



```

Question #3: For Democrats, what are the 5 words with the highest tf-idf values? What about for Republicans?

Democrats: affordable, glass, criminal, steagall, racism
Republicans: Marco, Ben, Megyn, maria, florida

```{r bigram model}
debate_bigrams <- debate_text %>% unnest_tokens(bigram, Text, token = "ngrams", n = 2)
debate_bigrams %>% count(bigram, sort = TRUE)




debate_bigrams <- debate_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>% 
  filter(!word2 %in% stop_words$word) %>% 
  unite(bigram, word1, word2, sep = " ")



bigram_count <- debate_bigrams %>% count(bigram, sort = TRUE) 

head(bigram_count, 20) %>% 
  ggplot(aes(x = reorder(bigram, n), y = n)) + 
  geom_bar(stat = "identity") + coord_flip()
```

Question #4: What are the 5 most frequent bigrams?

answer: Hillary Clinton, Wall street, american people, health care, President obama


```{r bigrams by party}

bigram_count_by_party <- debate_bigrams %>% count(Party, bigram, sort = TRUE) 

bigrams_tf_idf_top <- bigram_count_by_party %>%
  bind_tf_idf(bigram, Party, n) %>% 
  arrange(desc(tf_idf)) %>% 
  group_by(Party) %>%
  top_n(15)


ggplot(bigrams_tf_idf_top, aes(x = reorder(bigram, tf_idf), y = tf_idf, fill = Party)) +
  geom_bar(alpha = 0.8, stat = "identity", show.legend = FALSE) +
  facet_wrap(~Party, ncol = 2, scales = "free") + coord_flip() + 
  scale_fill_manual(values = c("Democratic" = "blue", "Republican" = "red"))
```

Question #5: For Democrats, what are the 3 bigrams with the highest tf-idf values? How about for Republicans?

answer: for democrats top 3 are senator sanders, affordable care, care act and for republicans the top 3 are eminent domain, free enterprise, trade center



